Optimizing Large Scale Surveys for Photometric Calibration, \\ The Euclid Dark Energy Mission as an Example

### 2011-07-05 Documentation/cross_cal_paper.tex
euclid: paper update
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@18954 03a3cea6-2c03-0410-ac9b-9271b1c66c29

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{holmes@mpia.de}

@@

\subsection{Sky Catalog}
We first create a catalog of fake sources, both stars and galaxies, based on realistic densities within the magnitude range $m_{min} < m < m_{max}$. We choose these limits based on the properties of single exposure of the current configuration of Euclid NIR Imaging Channel (NIP): $m_{min} = 17$, the approximate saturation limit, and $m_{max} = 22$, the 10$\sigma$ limit. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes distributed according to the following powerlaw:
\begin{displaymath}
\log_{10} \frac{dN}{d\Omega\,dm} = a + b\,m
\end{displaymath}
Where $a$ sets the total source density and $b_{star} = 0.25$ steller and galaxy populations reported in XX. We make no distinction between stars and galaxies in our simulations. The random number generator is given the same seed, so that the same simulated sky is produced each time the simulation is run. This ensures that the differences observed are not a result of random fluctuations. We relate source magnitudes, $m$, to source fluxes, $s$, simply by: $m = 22.5 - 2.5\log_{10}(s)$. We therefore do not consider the imaging instrument's absolute response.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{survey.png}
\end{center}
\caption{An example catalog of fake sources used within the simulations. The sources are shown at their position on a portion of the sky (left) and as a histogram of their magnitudes (right).
\label{fig:sky}}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{camera_image.png}
\end{center}
\caption{A single exposure of the sky with the pointing coordinates $(\alpha = 9.2^\circ, \beta = 2.2^\circ)$ and an orientation of $\theta = 175.4^\circ$. Left: Plot of the entire fake sky, with the instrument's field-of-view shown in black and the measured sources highlighted in red. Right: The distribution of the highlighted sources on the instrument's focal plane. \label{fig:camera}}
\end{figure}

\subsubsection{Catalog to Measured Counts Conversion}
For a measurement $(i)$ the counts recorded from a source ($k$) depends on true instrument response ($f(\vec{q}_{true})$), which is a function of focal plane position $(\vec{x_i})$, and the source's true flux ($s$):

\begin{displaymath}
c_i = f(\vec{x_i} | \vec{q}_{true}) . s_{k} + e_{i}
\end{displaymath}

\noindent{}An error, drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$, is also introduced to the measurement. In this analysis, we do not consider temporal variations in the instrument's response.

\subsubsection{Noise Model}
\label{sec:noise_model}
We assume that the Euclid exposures will be background limited and that, for systematic reasons, we will hit an upper limit on the signal-to-noise for bright sources. We assume that the uncertainty variance on the counts are:

\begin{displaymath}
\sigma_{{ass}}^{2} = \alpha^{2} + \eta^{2}\, c^{2} 
\end{displaymath}
\noindent{}where $\alpha$ and $\eta$ are both constants. In fact, we complicate the model by applying an extra term to the count's uncertainty variance, which we do not take into account in our analysis:

\begin{displaymath}
\sigma_{{act}}^{2} = (1 + \epsilon) \,\alpha^{2} + \eta^{2}\, c^{2} 
\end{displaymath}
\noindent{}where $\epsilon$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$. An example of the assumed compared to the true uncertainty variances is shown in Fig. \ref{fig:invvar}.

@@

\subsubsection{Instrument Response Model}
We simulate the instrument response $f(\vec{x_i} | \vec{q}_{true})$ variations on both large and small scales. We model large scale structure, such as the residual from a flat-field calibration source, as a second order polynomial. We model small scale variations in the instrument response with XX sine and cosine contributions. We intentionally do not for the instrument response with a model of the same form, instead we fit with a XX order polynomial only.

@@

\noindent{}We solve for the following \textit{incomplete} (see Section \ref{sec:noise_model}) model.
\begin{displaymath}
c_i = f(\vec{x_i} | \vec{q}) . s_{k} + e_{i}
\end{displaymath}
where $s_k$ is the true flux from star $k$ and our error is drawn from the Normal Distribution $e_{i} = N(e|0,{\sigma_i}^2)$.

@@

\subsubsection{Source Flux Refinement (\textit{s\_step})}
The flux estimates of each source, $s_k$, are considered individually. At each \textit{s\_step} we can refine a sources's flux estimate by minimizing the error function with the latest instrument response parameters:

@@

>>>>

\newcommand{\vdag}{(v)^\dagger}
\newcommand{\myemail}{holmes@mpia.de}
\newcommand{\true}{\mathrm{true}}

@@

\subsection{Sky Catalog}
We first create a catalog of fake sources, both stars and galaxies, based on realistic densities within the magnitude range $m_{min} < m < m_{max}$. We choose these limits based on the properties of single exposure of the current configuration of Euclid NIR Imaging Channel (NIP): $m_{min} = 17$, the approximate saturation limit, and $m_{max} = 22$, the 10$\sigma$ limit. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes distributed according to the following powerlaw:
\begin{displaymath}
\log_{10} \frac{dN}{d\Omega\,dm} = a + b\,m
\end{displaymath}
Where $a$ sets the total source density and $b_{star} = 0.25$ steller and galaxy populations reported in XX. We make no distinction between stars and galaxies in our simulations. The random number generator is given the same seed, so that the same simulated sky is produced each time the simulation is run. This ensures that the differences observed are not a result of random fluctuations. We relate source magnitudes, $m$, to source fluxes, $s$, simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies.

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{temp.png}
\end{center}
\caption{An example catalog of fake sources used within the simulations. The sources are shown at their position on a portion of the sky (left) and as a histogram of their magnitudes (right).
\label{fig:sky}}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{camera_image.png}
\end{center}
\caption{A single exposure of the fake sky. Left: Plot of the entire fake sky, with the instrument's field-of-view shown. Right: The resultant distribution of the highlighted sources on the instrument's focal plane. \label{fig:camera}}
\end{figure}

\subsubsection{Measurement Model}
For a measurement $i$ the counts recorded from a source $k$ depends on true instrument response $f_{\true}$, which is a function of focal plane position $\vec{x_i}$, and the source's true flux $s$:

\begin{displaymath}
c_i = f_{\true}(\vec{x_i}) \, s_{k, \true} + e_{i}
\end{displaymath}

\noindent{}An error, drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$, is also introduced to the measurement. In this analysis, we do not consider temporal variations in the instrument's response.

\subsubsection{Noise Model}
\label{sec:noise_model}
We assume that the Euclid exposures will be background limited and that, for systematic reasons, we will hit an upper limit on the signal-to-noise for bright sources. We complicate the model by applying an extra term to the count's uncertainty variance, which we do not take into account in our analysis:
\begin{displaymath}
\sigma_{i, \true}^{2} = (1 + \epsilon_k) \, \alpha^{2} + \eta^{2}\, [ f_\true(\vec{x_i}) \, s_{k, \true} ]^2
\end{displaymath}
\noindent{}where $\epsilon$ is a random number, in the range [0.0, $\epsilon_{max}$), generated for each measurement $i$.

We do not take into account this final term in our analysis and we therefoer assume that the uncertainty variance on the counts are:
\begin{displaymath}
\sigma_{{i}}^{2} = \alpha^{2} + \eta^{2}\, c^{2}_i
\end{displaymath}
\noindent{}where $\alpha$ and $\eta$ are both constants. In fact, we complicate the model by applying an extra term to the count's uncertainty variance, which we do not take into account in our analysis:

An example of the assumed compared to the true uncertainty variances is shown in Fig. \ref{fig:invvar}.

@@

\subsubsection{Instrument Response Model}
We simulate the instrument response $f_\true(\vec{x_i} | \vec{q}_{\true})$ variations on both large and small scales. We model large scale structure, such as the residual from a flat-field calibration source, as a second order polynomial. We model small scale variations in the instrument response with XX sine and cosine contributions. We intentionally do not for the instrument response with a model of the same form.

@@

\noindent{}We solve for the following \textit{incomplete} (see Section \ref{sec:noise_model}) model.
\begin{displaymath}
c_i = f(\vec{x_i} | \vec{q}) . s_{k} + e_{i}
\end{displaymath}
where $s_k$ is a model flux for source $k$ and the error $e_i$ is drawn from the Normal Distribution $N(e|0,{\sigma_i}^2)$. We model the instrument response as a XX order polynominal.

@@

\subsubsection{Source Flux Refinement (\textit{s\_step})}
The model fluxes of each source, $s_k$, are considered individually. At each \textit{s\_step} we can refine a sources's flux estimate by minimizing the error function with the latest instrument response parameters:

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{density_of_stars.png}
\end{center}

\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{epsilon_max.png}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{m_min.png}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{mag_at_10_sigma.png}
\end{center}
\end{figure}
